---
title: "RTransposer: example of usage (2)"
author: "Viciana, F; Paiva,D"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RTransposer: ejemplo de uso (2)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

It shows how to use [R](https://www.r-project.org/) and **RTransposer** package to import the example (2) provided by [Alter in the "Transposer" package repository](https://github.com/ICPSR/IDS) 

## Source file preparation

Downloading the data from github repository: ICPSR/IDS corresponding to example 2 (UG_example_2) and uncompress the excel file *UG_Example_2_v10.xlsx*

```{r}
library(RTransposer)
require(data.table) 
require(readxl)     
rm(list = ls())     

tmpLoc <- tempdir() ; print(tmpLoc)  # creation of a temporary folder for files used and produced in this script
## Download and uncompress
zipfile <-  paste0(tmpLoc,'/UG_example_2.zip')

download.file('https://github.com/ICPSR/IDS/raw/master/User_Guide/UG_example_2.zip', 
              destfile = zipfile )
unzip(zipfile,list = T)  

unzip(zipfile,files = 'UG_Example_2_v10.xlsx',  exdir =  tmpLoc )  
xfile <- paste0(tmpLoc,'/UG_Example_2_v10.xlsx')
file.remove(zipfile)

```
Data exrtraction from *parent* and *children* sheets, storage in a *data.table* type object and creation of some auxiliary/imput tables derived from these two source tables:
```{r}

as.data.table(read_xlsx(xfile, sheet = 'parents'))  -> parents
as.data.table(read_xlsx(xfile, sheet = 'children')) -> children

#------------------------------------
# Create INPUT tables derived from primaries (if necessary)  # in this case are 4, by sources (births, deaths, marriages) plus places
#------------------------------------
# places
names.loc  <- names(parents)[grep("loc",names(parents))]
names.loc  <- paste0('c(', paste(names.loc, collapse = ', '),', children$birthloc )')
places <- parents[, sort(unique(eval(parse(text=names.loc))))]

places_input<- data.table(
  place  = places,
  placeid = 1L:length(places)+100L,
  level  =  "Neighborhood",
  nested = 100L,
  relvar = "Neighborhood and Municipality"
)
  # append tow end register
places_input <-   rbind(places_input,
                        data.table(  place = c("California","Los Angeles"),
                                     placeid = c(1L,100L),
                                     level  =  c("State","Municipality"),
                                     nested = c(NA,1L),
                                     relvar = c("","Municipality and State")))

# marriages
parents[,.(parid,hfirst,hlast,wfirst,wlast,hocc,wocc,hid=parid*10+1,wid=parid*10+2,marriage,marrloc)]  -> marriages_input

# deaths
rbind(parents[,.(idi=parid*10+1,fname=hfirst,lname=hlast,death=hdeath,deathloc=hdeathloc)],
      parents[,.(idi=parid*10+2,fname=wfirst,lname=wlast,death=wdeath,deathloc=wdeathloc)]) -> deaths_input

# births
rbind(parents[,.(kidid=parid*10+1,fname=hfirst,lname=hlast,dadid=NA,momid=NA,birth=hbirth,birthloc=hbirthloc)],
      parents[,.(kidid=parid*10+2,fname=wfirst,lname=wlast,dadid=NA,momid=NA,birth=wbirth,birthloc=wbirthloc)],
      parents[,.(parid,lname=hlast)][
        children[,
                 .(fname=name,birth, birthloc, kidid= parid*1000+1:.N,
                   momid=parid*10+2,
                   dadid=parid*10+1),
                 keyby=.(parid)], on='parid'][,.(kidid,fname,lname,dadid,momid,birth,birthloc)]) -> births_input


```

If necessary, functions should be created to aid with the input, by transforming the original data stored in the source table; in this case it is created a function to convert geographic names (strings) into codes (integers):
```{r}
###-------------------------------------------------------------------------
###  auxiliary functions for ValueExpresion
###-------------------------------------------------------------------------
#' MunicipalityName2Code
#'    Input of a vector of municipality names and output of numeric codes
#'    contained in the table 
#'    @param mun.names vector con nombre
#'    @tab.codes  data.table with names and codes
MunicipalityName2Code <- function ( mun.names, tab.codes = places_input, 
                                    geo.names='places' ,geo.codes ='placeid' ) {
   tab.codes[[geo.codes]] -> codes
   names(codes) <- tab.codes[[geo.names]]
   return(codes[mun.names])
}

```

## Specifications of the import rules

The transformations applied to the variables in the source files are defined in a excel book with a specific structure (an example of this book can be found in the package's *inst/extdata/* folder) to obtain the records that compose the IDS. 

## Generating the database in IDS format

The *TRtransposer::tranposer* is applied, having as parameter the name of the [excel file](https://github.com/viciana/RTransposer/raw/master/inst/extdata/ERD_UG_Example_2_v10.xlsx) with the diferent transformations to be applied over the data tables, previously loaded in R's global enviroment (*.GlobalEnv).
```{r results="hide"}
##--------------------------------------------------------------------------
## Transposer with rules to import in  EntityRelationDefinition01.xlsx
transposer(file.definition = system.file("extdata", 
                                        "ERD_UG_Example_2_v10.xlsx", 
                                        package = "RTransposer") ) -> outcomes

```

By running the previous code a list is produced containing 5 *data.table* type objects with the IDS specific names and structure.

From this list it is possible to get the 5 tables in .CSV file, that can be easily read by statistical analysis base softwares.

The following exports these tables.

```{r}
### Statistical summary of outcomes::
print(paste('== INDIVIDUAL == ')) ;  print(outcomes$INDIVIDUAL[,.N,keyby=Type])
print(paste('== CONTEXT == ')); print(outcomes$CONTEXT[,.N,keyby=Type])
print(paste('== INDIV_INDIV == ')); print(outcomes$INDIV_INDIV[,.N,keyby=Relation])
print(paste('==  INDIV_CONTEXT == ')); print(outcomes$INDIV_CONTEXT[,.N,keyby=Relation])
print(paste('==  CONTEXT_CONTEXT == ')); print(outcomes$CONTEXT_CONTEXT[,.N,keyby=Relation])

## Storages the outcomes in CSV tables
for (i in  names(outcomes)) {
  write.csv(outcomes[[i]],file=paste0(tmpLoc,'/',i,'.CSV'),na='',row.names=FALSE)
}
print(tmpLoc)
dir(path = tmpLoc,pattern = '*.CSV')
```

Eliminates used temporary files
```{r results="hide", eval=FALSE}

file.remove(paste0(tmpLoc,'/',names(outcomes),'.CSV'))
file.remove(paste0(tmpLoc,'/','UG_Example_2_v10.xlsx'))
```

